---
title: "Load and clean data with R"
author: "Max Odsbjerg Pedersen"
date: "10/5/2020"
output: html_document
---

This notebook shows you how to load newspaper text data into a dataframe and how to extract text from a single newspaper into a text file.

First we load the library:

```{r, message=FALSE}
library(tidyverse)
```

# Import data

The code below imports data from Library Open Access Repository. Right now it is importing newspapers from 1840. You can change the year by finding a link to another years newspapers. You will find a list over years and their corresponding link containing that years newspaper data on the following link.

<https://datasprint2020.kb.dk/framework/>

When you have found your year of interest, copy the link and paste it in instead of the current link in the code below. Remember to change `artikler_1840` to a name corresponding to the year you chose.

```{r, message=FALSE}
artikler_1840 <- read_delim("https://loar.kb.dk/rest/bitstreams/ec134001-6c00-4658-95d5-76edd823adb2/retrieve",
  delim = ",",
  escape_backslash = TRUE,
  escape_double = FALSE)

```

## Inspecting the data

It is important before venturing into your data adventure to inspect and evaluate your data to make sure what you are dealing with and if you are dealing with the right thing.

In the code below we inspect the data we imported before with the `head()`-function:

```{r}
artikler_1840 %>% 
  head()
```

## Cleaning the data

The `editionId` column contains various metadata, among others the title of the newspapers. We want to extract the title from this column and store it in a new column. We do this by splitting on the space between the title and the other metadata

```{r, warning=FALSE}
artikler_1840 <- artikler_1840 %>% 
  separate(editionId, sep=" ", into = c("title"), remove = FALSE)
```

The dataframe now has a `title` column:

```{r}
artikler_1840 %>% 
  head()
```

## **Single newspaper text extraction**

Use the following method to extract all text of a specific newspaper from the dataframe into one string of text.

First, use the `count()` method to get a list of available newspaper title IDs.

```{r}
artikler_1840 %>% 
  count(title)
```

Select a newspaper using the filter function. We also use the `arrange()` to make sure that the text rows are chronological. Save this to a new dataframe

```{r}
artikler_1840 %>%
  arrange(sort_year_asc) %>% 
  filter(title == "faedrelandet1834") -> faedrelandet_1840
```

To combine just the text fields of all rows we select the column containing the text and next we collapse the rows into a single character element

```{r}
faedrelandet_1840_text <-faedrelandet_1840$fulltext_org %>% 
  paste(collapse = " ")
```

Next we save it to a local file:

```{r}
write_file(faedrelandet_1840_text, "faedrelandet_text.txt")
```
